### 2018-09-13 algorithm.tex

Taking the market clearing condition,
\begin{equation}
	R_{njt} = \sum_{m=1}^N
		d_{mnjt}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\end{equation}
\[
	R_{njt}\varrho_{njt}^{\theta} = \sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\]
\begin{equation}\label{eq:market_clearing}
	\varrho_{njt} = R_{njt}^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right]
	\right\}^{1/\theta}.
\end{equation}
Here both $R_{njt}$ and $\varrho_{njt}$ depend on equilibrium prices. But starting from a guess for $R$ (say, the free-trade equilibrium), we can easily compute a shadow price $\varrho$ supporting that allocation. 
\begin{enumerate}
	\item Set $R_{njt}^{(0)}$ and $P_{njt}^{(0)}$ to the autarky equilibrium value or other suitable starting value.
	\item Given $R_{njt}^{(k)}$ and $P_{njt}^{(k)}$, solve for $\varrho_{njt}^{(k+1)}$ using equation \eqref{eq:market_clearing}.
	\item Solve for $w_{njt}^{(k+1)}$ and $P_{njt}^{(k+1)}$ using \eqref{eq:wage} and \eqref{eq:price}.
	\item Calculate $R_{njt}^{(k+1)} = w_{njt}^{(k+1)}L_{njt}/\beta_j$. (This step sounds the most divergent, nothing ensures that relative sectoral revenue shares will remain close to equilibrium.)
	\item Go back to Step 2 and repeat until convergence.
\end{enumerate}

@@

The night labor share is a fixed point of
\[
E_{A}\frac 
	{\tilde w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\tilde w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*)}
= \lambda L_{nt}.
\]
\end{document}

>>>>

Taking the market clearing condition,
\begin{equation}\label{eq5}
	R_{njt} = \sum_{m=1}^N
		d_{mnjt}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\end{equation}
\[
	R_{njt}\varrho_{njt}^{\theta} = \sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\]
\begin{equation}\label{eq:market_clearing}
	\varrho_{njt} = R_{njt}^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right]
	\right\}^{1/\theta}.
\end{equation}
Here both $R_{njt}$ and $\varrho_{njt}$ depend on equilibrium prices. But starting from a guess for $R$ (say, the free-trade equilibrium), we can easily compute a shadow price $\varrho$ supporting that allocation. 
\begin{enumerate}
	\item Set $R_{njt}^{(0)}$ and $P_{njt}^{(0)}$ to the autarky equilibrium value or other suitable starting value.
	\item Given $R_{njt}^{(k)}$ and $P_{njt}^{(k)}$, solve for $\varrho_{njt}^{(k+1)}$ using equation \eqref{eq:market_clearing}.
	\item Solve for $w_{njt}^{(k+1)}$ and $P_{njt}^{(k+1)}$ using \eqref{eq:wage} and \eqref{eq:price}.
	\item Calculate $R_{njt}^{(k+1)} = w_{njt}^{(k+1)}L_{njt}/\beta_j$. (This step sounds the most divergent, nothing ensures that relative sectoral revenue shares will remain close to equilibrium.)
	\item Go back to Step 2 and repeat until convergence.
\end{enumerate}

@@

The night labor share is a fixed point of
\[
E_{A}\frac 
	{\tilde w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\tilde w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*)}
= \lambda L_{nt}.
\]
\subsection{Calibrating the CES model}
To calibrate the CES demand shifters $\nu$, we need data on how final expenditure is split across sectors in each country. We will recover this from the split of revenue, subtracting intermediate expenditure.

Let
\[
\Delta_{jt} = [d_{mnjt}] 
\]
denote the matrix of import shares (including the domestic market). We know that $\Delta_{jt}\mathbf 1=1$. We write equation \eqref{eq5} in matrix form as
\begin{equation}
	\tilde {\mathbf R}_{jt} = \tilde {\mathbf E}_{jt}\Delta_{jt},
\end{equation}
which we can invert
\begin{equation}
	\tilde {\mathbf E}_{jt} = \tilde {\mathbf R}_{jt}\Delta_{jt}^{-1},
	\end{equation}
and we can use $\tilde {\mathbf Y}_{jt}/\beta_j = \tilde{\mathbf R}_{jt}$. Once we have $E_{mjt}$, we subtract $\sum_k\gamma_{jk}R_{mkt}$ intermediate expenditure to get the final expenditure. Note that either of these expenditures may be negative in the data, in which case they are replaced by numerical zero.

Given a set of expenditure shares, we proceed as follows. First, we recover demand shifters and aggregate price index for the baseline country (US), using the observed sectoral prices there.
\[
\nu_{0jt} = e_{0jt} P_{0jt}^{\sigma-1} P_{0t}^{1-\sigma},
\]
where $P_{0t}$ is selected such that $\sum_j \nu_{0jt}=1$ for all $t$.

Second, for the tradable sectors, we recover sectoral prices $P_{njt}$ using the trade share formula. This expresses prices relative to the baseline country, but we already have the baseline prices.

Third, we recover tradable demand shifters using the price of the sector relative to the consumption aggregate obtained from PWT.
\[
\nu_{njt} = e_{njt} P_{njt}^{\sigma-1} (P_{0t}PWT_{nt})^{1-\sigma},
\]
for all $j<J$.

Fourth, for services, we calculate the demand shifter as 
\[
\nu_{nJt} = 1-\sum_{j\neq J}\nu_{njt},
\]
and construct its prices as
\[
P_{nJt} = (\nu_{nJt}/e_{nJt})^{1/(\sigma-1)} P_{0t}PWT_{nt}.
\]
This is the only step that has to be done separately for $\sigma=1$, although approximations may be valid. CHECK.
\end{document}
